---
title: "KGML gasmet data processing"
subtitle: "$CO_2$ $N_2O$ and $CH_4$"
author: "Jesse Bealsburg & Katherine Bohn"
date: today
format: pdf
editor: visual
execute:
  echo: false
  message: false
  warning: false
editor_options: 
  chunk_output_type: console
---

# Key points

-   $CO_2$ $N_2O$ and $CH_4$ converted from $\mathrm{ppm}$ to $\{\mathrm{elemental\ weight}\} \times  \{\mathrm{unit\ of\ area}\} \times \ \{\mathrm{duration}\}$  like $\mathrm{g} \cdot \mathrm{m}^2 \cdot \mathrm{day}^{-1}$

-   Most $CO_2$ fluxes are positive = good.

    -   105 of the 2836 predict positive $CO_2$ emissions

-   $N_2O$ and $CH_4$ fluxes are small and normally distributed around zero = concerning

-   Kathryn independtly processed gasmet data. This data was brought in for comparison and the distribution of $N_2O$ and $CO_2$ remained = further investigation and discussion required

    -   Is there a calibration issue for $N_2O$ and $CH_4$?

    -   Do we expect most N2O and CH4 data to be useless?

```{r, include=FALSE}
# I am importing the rmarkdown file authored by Jake Kundert that I modified, exporting it as a R script and running that rscript to populate my environment, then removing most of the objects in the environment except the ones I am keeping

knitr::purl("kgml_gasmet.fluxcalc.Rmd", output = "kgml_gasmet.fluxcalc_codeOutput.R")
source("kgml_gasmet.fluxcalc_codeOutput.R")

```

```{r}
# this is Jesse's outputs
coefs_ch4.sum %>% 
  # glimpse()
  select(experiment, site, plot, date, ch4.e.gmd) -> ch4_dat

coefs_n2o.sum %>% 
  # glimpse()
  select(experiment, site, plot, date, n2o.e.gmd) -> n2o_dat

coefs.sum %>% 
  select(experiment, site, plot, date, co2.e.gmd) -> co2_dat

# Define objects to keep
keep <- c("coefs_ch4.sum","ch4_dat",
          "coefs_n2o.sum", "n2o_dat", 
          "coefs.sum", "co2_dat")

# Remove everything except those
rm(list = setdiff(ls(), keep))

```

```{r}

# joining data for export to kgml google sheet and cleaning up naming

n2o_dat %>% 
  left_join(co2_dat) %>% 
  left_join(ch4_dat) %>% 
  mutate(experiment=tolower(experiment)) %>% 
  mutate(experiment = fct_recode(experiment,
                                 "eddyflux" = "eddyfux",
                                 "eddyflux" = "eddyfluedddyflux")) %>% 
  mutate(site = tolower(site)) %>% 
  filter(!grepl("w", plot)) %>% 
  filter(grepl("^\\d+$", plot)) -> dat1
  
# dat1 %>%
#   write.csv("kgml_data_daily_postProcess_1Apr2025.csv",
#             row.names = F)
```

# Visualization

## Jesse

```{r}
## CO2

dat1 %>% 
  ggplot(aes(co2.e.gmd)) +
  stat_bin()

# dat1 %>% 
#   filter(co2.e.kghyr<0) %>%
#   # View()
#   tally()

# 105 of the 2836 data points have negative slopes, there is no obvious reason.


## N2O
dat1 %>% 
  ggplot(aes(n2o.e.gmd)) +
  stat_bin()


# dat1 %>% 
#   filter(n2o.e.kghyr<0) %>%
#   group_by(experiment) %>% 
#   tally()

# 1321 files have negative slopes of the 2836 data points
# 42-51% of each experiment has negative vs positive flux

dat1 %>% 
  filter(n2o.e.gmd<0) %>%
  group_by(date) %>%
  tally() %>% 
  arrange(desc(n)) %>% 
  dplyr::rename(negSlopeTally = n) -> dat_negSlope


dat1 %>% 
  group_by(experiment,date) %>%
  tally() %>% 
  arrange(desc(n)) %>% 
  dplyr::rename(totalSlopeTally = n) %>% 
  left_join(dat_negSlope) %>% 
  mutate(negFluxPercent = negSlopeTally/totalSlopeTally*100) %>% 
  arrange(desc(negFluxPercent)) %>% 
  # dplyr::rename(
  #   "total fluxes predicted (n)" = "totalSlopeTally",
  #   "negative fluxes predicted" = "negSlopeTally",
  #   "percent of fluxes that are negative" = "negFluxPercent"
  # ) %>% 
  print(n=10)


# CH4

dat1 %>% 
  ggplot(aes(ch4.e.gmd)) +
  stat_bin()

dat1 %>% 
  filter(ch4.e.gmd<0) %>%
  group_by(date) %>%
  tally() %>% 
  arrange(desc(n)) %>% 
  dplyr::rename(negSlopeTally = n) -> dat_negSlope

dat1 %>% 
  group_by(experiment,date) %>%
  tally() %>% 
  arrange(desc(n)) %>% 
  dplyr::rename(totalSlopeTally = n) %>% 
  left_join(dat_negSlope) %>% 
  mutate(negFluxPercent = negSlopeTally/totalSlopeTally*100) %>% 
  arrange(desc(negFluxPercent)) %>% 
  # dplyr::rename(
  #   "total fluxes predicted (n)" = "totalSlopeTally",
  #   "negative fluxes predicted" = "negSlopeTally",
  #   "percent of fluxes that are negative" = "negFluxPercent"
  # ) %>% 
  print(n=10)

```

## Kathryn's 

I'd expect her process to be more robust than mine. Just ATAT and Eddyflux

```{r}
read.csv("atat_GHG_coefs_2022thru2023 - atat_GHG_coefs_2022thru2023.csv") %>% 
  mutate(experiment="atat") %>% 
  select(
    experiment,site,plot,date,
    slope_CO2_ugcm3,
    slope_N2O_ugcm3,
    slope_CH4_ugcm3,
    co2.e.gmhr,
    n2o.e.gmhr,
    ch4.e.gmhr
  )-> dat_kat_atat

read.csv("eddyflux_GHG_coefs_2024.csv") %>% 
    select(
    experiment,site,plot,date,
    slope_CO2_ugcm3,
    slope_N2O_ugcm3,
    slope_CH4_ugcm3,
    co2.e.gmhr,
    n2o.e.gmhr,
    ch4.e.gmhr
  ) -> dat_kat_eddyflux

# dat_kat_atat %>% 
#   bind_rows(dat_kat_eddyflux) %>% 
#   write.csv(
#     "kgml_KathrynPhillips_atatEddyflux_4Apr2025.csv",
#     row.names = F
#             )


dat_kat_eddyflux %>% 
  ggplot(aes(co2.e.gmhr)) + 
  stat_bin()

dat_kat_eddyflux %>% 
  ggplot(aes(n2o.e.gmhr)) + 
  stat_bin()

dat_kat_eddyflux %>% 
  ggplot(aes(ch4.e.gmhr)) + 
  stat_bin()

dat_kat_eddyflux %>% 
  ggplot(aes(ch4.e.gmhr)) + 
  stat_bin()

```
