---
title: "KGML gasmet data processing"
subtitle: "CO2 and N2O"
author: "Jesse Bealsburg"
date: today
format: pdf
editor: visual
execute:
  echo: false
  message: false
  warning: false
editor_options: 
  chunk_output_type: console
---

# Key points

-   Converted ppm of CO2 to kg C ha yr and N2O to kg N ha yr

    -   fitting lines, extracting slope

-   Most CO2 fluxes are positive, which is what it should be.

    -   105 of the 2836 predict positive CO2 emissions

-   \~50% of N2O fluxes are positive.

Yufeng wants N2O data converted from the ppm unit to an amount of N released over an area and amount of time. Here I convert N2O ppm to kg N ha over a year by modifying the script used by Jake Kundert to convert CO2 to kg C ha yr.

I am importing the rmarkdown file authored by Jake Kundert that I modified, exporting it as a R script and running that rscript to populate my environment, then removing most of the objects in the environment except the ones I am keeping

```{r, include=FALSE}
knitr::purl("kgml_gasmet.fluxcalc.Rmd", output = "kgml_gasmet.fluxcalc_codeOutput.R")
source("kgml_gasmet.fluxcalc_codeOutput.R")
```

```{r}
coefs_n2o.sum %>% 
  select(experiment, site, plot, date, n2o.e.kghyr) -> n2o_dat

coefs.sum %>% 
  select(experiment, site, plot, date, co2.e.kghyr) -> co2_dat

# Define objects to keep
keep <- c("coefs_n2o.sum", "n2o_dat", 
          "coefs.sum", "co2_dat")

# Remove everything except those
rm(list = setdiff(ls(), keep))

```

This data has already undergone some QAQC. I will simply join and export here. In general, CO2 data is a good indicator of the quality of the N2O data. If the standard error was abnormally high for CO2, that same sample likely has unreliable N2O data.

```{r}

# joining data for export and cleaning up naming

n2o_dat %>% 
  left_join(co2_dat) %>% 
  mutate(experiment=tolower(experiment)) %>% 
  mutate(experiment = fct_recode(experiment,
                                 "eddyflux" = "eddyfux",
                                 "eddyflux" = "eddyfluedddyflux")) %>% 
  mutate(site = tolower(site)) %>% 
  filter(!grepl("w", plot)) %>% 
  filter(grepl("^\\d+$", plot)) -> dat1
  
# dat1 %>% 
#   write.csv("kgml_data_daily_postProcess_1Apr2025.csv",
#             row.names = F)
```

# Visualization

```{r}
## CO2

dat1 %>% 
  ggplot(aes(co2.e.kghyr)) +
  stat_bin()

# dat1 %>% 
#   filter(co2.e.kghyr<0) %>%
#   # View()
#   tally()

# 105 of the 2836 data points have negative slopes, there is no obvious reason.


## N2O
dat1 %>% 
  ggplot(aes(n2o.e.kghyr)) +
  stat_bin()


# dat1 %>% 
#   filter(n2o.e.kghyr<0) %>%
#   group_by(experiment) %>% 
#   tally()

# 1321 files have negative slopes of the 2836 data points
# 42-51% of each experiment has negative vs positive flux

dat1 %>% 
  filter(n2o.e.kghyr<0) %>%
  group_by(date) %>%
  tally() %>% 
  arrange(desc(n)) %>% 
  dplyr::rename(negSlopeTally = n) -> dat_negSlope


dat1 %>% 
  group_by(experiment,date) %>%
  tally() %>% 
  arrange(desc(n)) %>% 
  dplyr::rename(totalSlopeTally = n) %>% 
  left_join(dat_negSlope) %>% 
  mutate(negFluxPercent = negSlopeTally/totalSlopeTally*100) %>% 
  arrange(desc(negFluxPercent)) %>% 
  # dplyr::rename(
  #   "total fluxes predicted (n)" = "totalSlopeTally",
  #   "negative fluxes predicted" = "negSlopeTally",
  #   "percent of fluxes that are negative" = "negFluxPercent"
  # ) %>% 
  print()

```
