---
title: "KGML notebook"
date: today
format: pdf
editor: visual
include: false
---

# Conversions

solving for the conversion factor from plot yield to kilograms per hectare

$$ y  = mx $$

where $y$ is grain yield in desired unit, likely kg ha

where $x$ is grain yield in grams per harvest area

where $m$ is the conversion factor

## Fert 

### corn and soy yields

```{r, message=FALSE}
library(tidyverse)
library(measurements)

## corn and soy grain plot yields to kg ha

conv_unit(30,"inch","m") * 2 * 3
# assumes 30 inch row spacing, converting that meter basis. each row experiences 30in row spacing. so 2 rows at 30 inch row spacing for 3 meters experiences 3 meters long and 2*30inch row spacing

# 4.572 square meter area

# corn grain yield assumed reported in grams

conv_unit(10000,"m2","hectare")

1/4.572*10000/1000


```

$$
\begin{aligned}
\frac{\text{grain yield (g)}}{\text{plot}} \times \frac{\text{plot}}{4.572 \, m^2} \times \frac{10000 \, m^2}{1 \, \text{hectare}} \times \frac{1 \, \text{kg}}{1000 \, \text{g}} = 2.187227 \\
\text{plot yield (g plot}^{-1}\text{)} \times 2.187 = \text{plot yield (kg ha}^{-1}\text{)} \\
y = 2.187 \, x 
\end{aligned}
$$

### IWG yields

```{r, message=FALSE}

## iwg grain plot yields to kg ha

conv_unit(36*36,"inch2","m2")*2

# 1.672255 square meter area

# iwg  grain yield assumed reported in grams

conv_unit(10000,"m2","hectare")

1/1.672255*10000/1000



```

$$
\begin{aligned}
\frac{\text{iwg weight }g}{\text{sample area}} \times \frac{\text{sample area}}{1.672 \, m^2} \times \frac{10000 m^2}{1 hectare} \times \frac{1 kg}{1000 g} = 5.98\\
y = 5.98x
\end{aligned}
$$

# Issues with data

## ORG 

-   outliers for corn and soybean yields, likely remove from data

-   IWG yields have some values above 2000 kg ha 😕

```{r, include=TRUE, echo=FALSE}

library(tidyverse)

read.csv("jesse_KGML_Dataset.xlsx - Data.csv") %>% 
  filter(project=="ORG") %>% 
  select(crop, location, year, seedyld) %>% 
  mutate(type = "actual") -> dat_actual

tibble(
  crop = c("iwg","corn","soybean"),
  seedyld = c(500,12000,4000),
  type = c(rep("predicted_Jesse",3))) -> dat_predicted
  
dat_actual %>% 
  ggplot(aes(crop,seedyld)) +
  geom_point(aes(col=type)) +
  geom_point(
    data = dat_predicted,
    aes(col = type),
    size=3,
    position = position_dodge(.5)
  ) 

```

| Crop       | Jesse            | ChatGPT          | Google           |
|------------|------------------|------------------|------------------|
|            | seed yield kg ha | seed yield kg ha | seed yield kg ha |
| corn       | 12000            | 7000-12000       | 7000             |
| soybean    | 4000             | 2000-45000       | 4000-11000       |
| wheatgrass | 500              | 600              | 600              |

: estimated seed yield by crop in kg ha. I used a conversion from bushels acre yields + test weight to kg ha

### Closer look at IWG

```{r, include=TRUE, echo=FALSE}
dat_actual %>% 
  filter(crop=="iwg") %>% 
  ggplot(aes(crop,seedyld)) +
  geom_point(aes(col=type)) +
  geom_point(
    data = subset(dat_predicted, crop=="iwg"),
    aes(col = type),
    size=3,
    position = position_dodge(.5)
  ) +
  geom_boxplot(fill=NA, 
               width=.1) +
  labs(
    y="seed yield kg ha"
  )

```

-   kernza grain yields above 2000 kg ha are suspicious, and are noted as a boxplot to be outliers
